We used a questionnaire in order to understand the current experiences of
supercomputer centers with respect to interacting with their electricity
service providers. We restricted the analysis to sites in the United States
since the results of the survey and practices of demand response is highly
correlated and driven by energy policies in the country. 
\cite{torriti_demand_2010}
[Tor10].

Nineteen Top100 List sized sites in the United States were targeted for the
questionnaire. Eleven sites responded (Oak Ridge National Laboratory, 
Lawrence Livermore National Laboratory, 
Argonne National Laboratory, 
Los Alamos National Laboratory, LBNL, 
Wright Patterson Air Force Base,
National Oceanic Atmospheric Administration, 
National Center for Supercomputing Applications, 
San Diego Supercomputing Center, 
Purdue University and Intel Corporation). Eight sites didn't respond (
National Center for Atmospheric Research,
IBM Corporation, 
National Energy Technology Laboratory, Indiana University, 
Texas Advanced Computing Center, 
Sandia Natinal Laboratory, 
National Renewable Energy Laboratory, 
National Aeronautics and Space Administration). The questionnaire was
sent to a sample that was not randomly selected. It was sent to those sites
where it was relatively easy to identify an individual based on membership
within the EE HPC WG. The sample is more representative of Top50 sized sites
(1 Top50 sized site was not in the sample and 60{\%} (9/15) of the sample
responded). Only 4 additional sites were sampled from the Top51-Top100 List
and, of those, 2 responded (Intel and National Oceanic and Atmospheric Administration).

The total power load as well as the intra-hour fluctuation of these sites
varied significantly. There were four sites with total power load greater
than 10MW, two sites with \textasciitilde 5MW total power load and five
sites with less than 2MW total power load. We chose less than 3MW intra hour
variability as the bottom of the scale because we assumed that the
electrical service providers would not be affected by that magnitude of
fluctuation. For those with total power load greater than 10MW, the
intra-hour fluctuation varied from less than 3MW to 8MW. One of
\textasciitilde 5MW sites said that they experienced 4MW variability. The
rest of the sites were all less than 3MW. Most of the intra-hour variability
was due to preventative maintenance.


\begin{table}[htbp]

\begin{center}
\caption{Caption Number 1}
\begin{tabular}{|p{65pt}|l|l|}
\hline
\textbf{Total Load}&
\textbf{Variability}&
\textbf{Frequency} \\
\hline
16-17MW&
5MW&
weekly \\
\hline
13-14MW&
8MW&
monthly \\
\hline
10-11MW&
Less than 3MW&
weekly \\
\hline
10-11MW&
7MW&
weekly \\
\hline
4-5MW&
Less than 3MW&
weekly \\
\hline
4-5MW&
4MW&
weekly \\
\hline
1-2MW&
Less than 3MW&
weekly \\
\hline
1-2MW&
140kW&
daily \\
\hline
1-2MW&
Less than 3MW&
yearly \\
\hline
1-2MW&
200kW or less&
daily \\
\hline
1-2MW&
Less than 3MW&
daily \\
\hline
\end{tabular}
\label{tab1}
\end{center}
\end{table}

We asked if the supercomputer centers had talked to their electric service
providers about programs and methods used to balance the grid supply and
demand of electricity. About half of them have had some discussion, but it
has mostly been limited to demand side and not supply-side programs.


\begin{table}[htbp]

\begin{center}
\caption{Caption Number 2}
\begin{tabular}{|p{230pt}|l|}
\hline
\textbf{Discussions with Electricity Providers}&
{\%} Answered Yes \\
\hline
\textbf{Programs}&
~ \\
\hline
Shedding load during peak demand&
54 \\
\hline
Responding to pricing incentive programs&
45 \\
\hline
Shifting load during peak demand&
36 \\
\hline
\textbf{Methods}&
~ \\
\hline
Enabling use of renewables&
36 \\
\hline
Congestion, Regulation, Frequency Response&
18 \\
\hline
Contributing to electrical grid storage&
10 \\
\hline
\end{tabular}
\label{tab2}
\end{center}
\end{table}

More than half of the respondents are not interested in shedding or shifting
load during peak demand. There is some indication that this low interest is
primarily due to the lack of a clear business case. For the sites where
there is interest, shifting is more attractive than shedding load. The San Diego Supercomputing Center (SDSC) is
an exception to this trend, but because of a site-wide program. ``University of California San Diego
generates 30-35MW of power yet still imports 5-10MW. As a large generation
source the utility providers see the campus as a highly attractive partner
for offloading grid stress. Automatic load shedding is being
explored/deployed today.''

Responding to pricing incentive programs is also not considered interesting,
although the reasons for this low interest may be organizational. Several
open-ended comments revealed that pricing is fixed and/or done by another
organization at the site level and outside of their immediate control.

Eighty percent of the respondents have not had discussions with their
electricity service providers about congestion, regulation and frequency
response. Los Alamos National Laboratory (LANL) is one of two who have had discussions and who commented that
they are ``learning about the process'' and that it is ``outside of [their]

visibility or control''.

There were been many more respondents who have had discussions with their
electricity service providers about enabling the use of renewables; 36{\%}
have already had discussions and more than half are interested in further
and/or future discussions. SDSC already has a site-wide program; ``the
campus has a large fuel cell (2.5$+$ MW) and works with the utility with
renewables.'' Other responses suggest that the interest is at the site level
and not unique to the supercomputer center.

An open-ended question was posed as to whether or not there was information
either requested of the supercomputer sites by their providers or,
conversely, requested of the providers by the sites. In both cases, well
over 75{\%} of the respondents answered no. Lawrence Livermore National Laboratory (LLNL) and LANL were the
exceptions. LLNL is ``working on obtaining additional data from them and a
means of sharing data between them and us'' and has been requested to
provide ``additional detailed forecasting and ultimately real time data.''
LANL has also been requested to provide ``power projections, hour by hour,
for at least a day in advance'' and, perhaps as a consequence, would like to
have more information on ``sensitivity of power distribution grid to rapid
transients (random daily step changes of 10 MW up or down within a single AC
cycle).''

Given the low levels of current engagement between the electricity service
providers and the supercomputer centers, it is not surprising that none of
the supercomputer centers are currently using any power management
strategies to respond to grid requests by their electrical service
providers. SDSC's \textit{supercomputer center} is not an exception, but they did respond that their
entire ``campus is leveraging parallel electrical distribution to trigger
diesel generators and other back-up resources to respond to to grid and
non-grid requests.''

We tried to evaluate if power management strategies will be considered
relevant and effective for grid integration at some point in the future. Two
questions were asked; is there interest in using the strategies and what
impact did they think that the strategies would have. When combining
interest and impact, the results showed that power capping, shutdown, and
job scheduling were both high interest and impact. Load migration, back-up
scheduling, fine grained power management and thermal management were medium
interest and impact. Lighting control and back-up resources were low
interest and impact.


\begin{table}[htbp]

\begin{center}
\caption{Caption Number 3}
\begin{tabular}{|p{299pt}|}
\hline
\textbf{HPC strategies for responding to Electricity Provider requests (listed from highest to lowest interest }$+$\textbf{ impact)} \\
\hline
Course grained power management \\
\hline
Facility shutdown \\
\hline
Job scheduling \\
\hline
Load migration \\
\hline
Re-scheduling back-ups \\
\hline
Fine grained power management \\
\hline
Temperature control beyond ASHRAE limits \\
\hline
Turn off lighting \\
\hline
Use back-up resources (e.g., generators) \\
\hline
\end{tabular}
\label{tab3}
\end{center}
\end{table}
